Model estimation

### 2024-11-23 modelestimate.Rmd
nullmodelcorrection

First, I estimate a null model with only reciprocity and degree. The
maximum convergence of the model is **NUMBER** The GOF on the Indegree
Distribution is **NUMBER** To have all GOF plots compiled in one place,
see here: *Link GOF page* This indicates that the model is a bad fit for
the data, which was to be expected since we ran a null model without any
network effects.

**CHANGE** Both reciprocity and density are significant for an alpha of
**NUMBER**.

@@

```{r effects M0}

myeff_M0 <- includeEffects(myeff_M0, isolateNet, inPop, outAct)

<!-- # Null model: just myeff (just degree and reciprocity by default) # model0  #for this when you create Siena Data, do not add the gender variable -->

@@

```{r estimate M0, eval=FALSE}
# estimate the model 
Model0 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, returnDeps = TRUE)
Model0
# maximum convergence >.25
# run again:
Model0.1 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, prevAns= Model0, returnDeps = TRUE)
Model0.1

#fsave(Model0, file ="estimated_M0")
#fsave(Model0.1, file = "estimated_M0.1")

@@

Model0.1 <- fload("./data/processed/estimated_M0.1_20241123.rda")
Model0.1

@@

```{r GOF M0}
GOF_M0.1 <- sienaGOF(Model0.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M0.1, main = 'GOF - "Null" Model 0 \n Indegree Distribution')
```

@@

**CHANGE** "It has a convergence of **NUMBER** and a Goodness of Fit on
the indegree distribution of **NUMBER**."

@@

@@

**CHANGE** "This model has a convergence of **0.23** and a Goodness of
Fit on the indegree distribution of **0.184**. For the structural
network statistics we find similar results, in terms of estimate size
and significance, as in Model 1. However, the outdegree activity
parameter almost reaches significance (**-1.7135484**).

@@

@@

This model is partly similar to Model 2 but additionally holds the
network statistics egoX and altX for the main effect of department, to
test if an alter's department affects an ego's tendency to initiate
collaborations and to see if department affects an individual’s
attractiveness as a collaborator.

**CHANGE** "This model has a convergence of **0.23** and a Goodness of
Fit on the indegree distribution of **0.184**. For the structural
network statistics we find similar results, in terms of estimate size
and significance, as in Model 1. However, the outdegree activity
parameter almost reaches significance (**-1.7135484**).

@@

@@

# if maximum convergence >.25: run again
#Model3.1 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, prevAns= Model3, returnDeps = TRUE)
#Model3.1

@@

#Model3.1 <- fload ("./data/processed/estimated_M3.1_20241123.rda")
#Model3.1
```

@@

```{r GOF M3}
GOF_M3 <- sienaGOF(Model3, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M3, main = "GOF - Model 3 \n Indegree Distribution")
```

@@

**CHANGE!!!** "This model has a convergence ratio of **0.23** and a a
Goodness of Fit on the indegree distribution of **0.64**. In this model,
the parameters for density and reciprocity are still significant, with
values of **-1.98** (se = **0.71**) and **2.59** (se = **0.38**)
respectively, meaning that actors prefer sparse networks, but do prefer
to send ties to alters who have sent ties to them.

@@

@@

myeff_M3 <- includeEffects(myeff_M3, isolateNet, inPop, outAct) # model 1

@@

fsave(Model4, file ="estimated_M4")

>>>>

First, I estimate a null model with only reciprocity and degree.

The maximum convergence ratio of the model is 0.084. The GOF on the
Indegree Distribution is zero**.** This indicates that the model is a
very bad fit for the data. This is not suprising since we ran a null
model without any network effects.ffects.

For all GOF plots compiled in one place, see here:
<https://treesdude.github.io/labjournal/finalGOF.html>

CHANGE Both reciprocity and density are significant for an alpha of
**NUMBER**.

Estimates and standard errors

Rate parameters:

0\. Rate parameter 2.117 ( 0.3561)

Other parameters:

1\. eval: outdegree (density) -2.6388 ( 0.1460) convergence t ratio:
0.0182

2\. eval: reciprocity 1.8035 ( 0.4343) convergence t ratio: -0.0668

@@

Null model: none.

@@

```{r estimate M0, eval=FALSE}
# estimate the model 
Model0 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, returnDeps = TRUE)
Model0

fsave(Model0, file ="estimated_M0")

@@

@@

```{r GOF M0}
GOF_M0 <- sienaGOF(Model0, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M0, main = 'GOF - "Null" Model 0 \n Indegree Distribution')
```

@@

"The model has a maximum convergence ratio of **0.17** and a Goodness of
Fit on the indegree distribution of **.002.**"

@@

Estimates and standard errors

Rate parameters:

0\. Rate parameter 3.3566 ( 0.5773)

Other parameters:

1\. eval: outdegree (density) -2.0862 ( 0.7050)

2\. eval: reciprocity 1.0894 ( 0.4144)

3\. eval: indegree - popularity 0.2549 ( 0.0439)

4\. eval: outdegree - activity -0.0936 ( 0.1301)

5\. eval: network-isolate 4.9645 ( 1.1146)

convergence t ratio:

```         
1
0.0232
```

```         
2
0.0467
```

```         
3
-0.0195
```

```         
4
-0.0423
```

```         
5
-0.061
```

```         
 -0.0617
```

@@

```         
Estimates, standard errors and convergence t-ratios                                                
                                            Estimate   Standard   Convergence
                                                        Error      t-ratio     
Rate parameters:    
0       Rate parameter                       2.9204  ( 0.4617   )

Other parameters:    
1. eval outdegree (density)                 -1.8628  ( 1.1060   )    0.0703      
2. eval reciprocity                          1.3341  ( 0.4543   )   -0.0152      
3. eval indegree - popularity                0.2209  ( 0.0470   )   -0.0083      
4. eval outdegree - activity                -0.3017  ( 0.2287   )    0.0806      
5. eval network-isolate                      4.7507  ( 1.5194   )   -0.0379      
6. eval gender_RS alter                      0.3254  ( 0.2701   )    0.0411      
7. eval gender_RS ego                       -1.0642  ( 0.6055   )   -0.0400      
8. eval same gender_RS                       0.6637  ( 0.2685   )    0.0308      
9. eval transitive triplets diff. gender_RS  2.1787  ( 0.6027   )    0.0416   
```

The maximum convergence ratio for model 2 is **0.132** and the Goodness
of Fit on the indegree distribution is **0.096**.

**CHANGE!!**

"For the structural network statistics we find similar results, in terms
of estimate size and significance, as in Model 1. However, the outdegree
activity parameter almost reaches significance (**-1.7135484**).

@@

egoX and altX: model whether male or female scholars are more likely to
initiate or receive ties.

**sameX (Gender)**: Tests whether scholars prefer same-gender
collaborators (direct measure of gender homophily).\
model the general tendency for scholars to collaborate with same-gender
peers.

**Confirmation**:

-   Significant and positive coefficient for `sameX` in the RSiena
    output would confirm that gender homophily exists.

-   A non-significant or negative coefficient would disconfirm this
    hypothesis.

**Confirmation**:

-   Significant and positive coefficient for `sameX` in the RSiena
    output would confirm that gender homophily exists.

-   A non-significant or negative coefficient would disconfirm this
    hypothesis.

Hypothesis about structural effects

-   **Network Statistics**:

    -   **sameX (Gender)**: Tests for gender homophily directly.

    -   Structural statistics like **reciprocity**, **inPop**,
        **outAct**, and **isolateNet** are control variables to account
        for network-level dynamics.

-   **Confirmation**:

    -   A significant `sameX` coefficient, even after including
        structural controls, would confirm that gender homophily exists
        at the dyadic level.

    -   If `sameX` becomes non-significant after controls, it suggests
        structural factors, not individual preferences, explain gender
        clustering.

@@

This model is partly similar to Model 2 but additionally holds the
network statistics egoX and altX for the main effect of department, to
control for differences in collaboration patterns between the two
departments.

Model 3 has a maximum convergence ratio of **0.182** and a Goodness of
Fit on the indegree distribution of **0.108**.

**CHANGE** !!!!

For the structural network statistics we find similar results, in terms
of estimate size and significance, as in Model 1. However, the outdegree
activity parameter almost reaches significance (**-1.7135484**).

@@

Estimates and standard errors

Rate parameters:

0\. Rate parameter 3.1013 ( 0.5186)

Other parameters:

1\. eval: outdegree (density) -4.3233 ( 1.0871)

2\. eval: reciprocity 1.0725 ( 0.4174)

3\. eval: indegree - popularity 0.2122 ( 0.0517)

4\. eval: outdegree - activity -0.3364 ( 0.2182)

5\. eval: network-isolate 4.2006 ( 1.3007)

6\. eval: gender_RS alter 0.3356 ( 0.2659)

7\. eval: gender_RS ego -1.0066 ( 0.5121)

8\. eval: same gender_RS 0.6632 ( 0.2623)

9\. eval: transitive triplets diff. gender_RS 1.5479 ( 0.5403)

10\. eval: departm_RS alter 0.4141 ( 0.7546)

11\. eval: departm_RS ego -1.1573 ( 0.9148)

12\. eval: same departm_RS 2.7240 ( 0.7277)

-   **Network Statistics**:

    -   **sameX (Gender)**: Compute and compare the size and
        significance of the coefficient for each department. This
        requires separate models for each department or interaction
        terms that allow the effect to vary by department.

-   **Confirmation**:

    -   A significantly stronger positive coefficient for `sameX` in
        Political Science compared to Sociology would confirm the
        hypothesis.

    -   Similar coefficients or reversed patterns would disconfirm the
        hypothesis.

@@

#if maximum convergence >.25: run again
Model3.1 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, prevAns= Model3, returnDeps = TRUE)
Model3.1

@@

Model3.1 <- fload ("./data/processed/estimated_M3.1_20241123.rda")
Model3.1
```

@@

```{r GOF M3}
GOF_M3.1 <- sienaGOF(Model3.1 , IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M3.1, main = "GOF - Model 3 \n Indegree Distribution")
```

@@

"The last model has a maximum convergence ratio of **0.23** and a a
Goodness of Fit on the indegree distribution of **0.197**.

**CHANGE!!!**

"In this model, the parameters for density and reciprocity are still
significant, with values of **-1.98** (se = **0.71**) and **2.59** (se =
**0.38**) respectively, meaning that actors prefer sparse networks, but
do prefer to send ties to alters who have sent ties to them.

@@

About Hypothesis 4

modeling how the department of the actor (Sociology or Political
Science) influences their collaboration choices in terms of gender.

**Interaction1 = "departm_RS"**: This allows you to test how department
(Sociology vs Political Science) influences the probability of forming
collaborations, potentially differing by gender. This means that gender
homophily (whether a person collaborates more often with same-gender
peers) might be different in the two departments.

allows you to explore how **department** influences the likelihood of
**same-gender** collaborations (`sameX`), which is central to your
research question. If the **department variable** is acting as an
important covariate influencing collaboration patterns, this line
captures that effect.

testing how being in a particular department influences the likelihood
of collaborating with someone of the same gender. In other words, it
checks if the department modifies the strength of gender homophily
(`sameX`).

**Network Statistics**:

-   **egoX (h-index)**: Examines how the h-index of the focal scholar
    affects their collaboration behavior.

-   **altX (h-index)**: Examines how the h-index of potential
    collaborators affects their attractiveness as partners.

    Add `egoX` and `altX` effects for `HIndex_RS` to test if prestigious
    scholars are more likely to initiate or receive ties.

-   **inPopX (h-index)**: Assesses whether prestige increases the
    likelihood of being sought out as a collaborator.

**Confirmation**:

-   A significant negative interaction term between `sameX` and h-index
    statistics would confirm that prestige reduces the strength of
    gender homophily.

-   If h-index variables do not significantly influence `sameX`, this
    would disconfirm H3.

@@

myeff_M4 <- includeEffects(myeff_M4, isolateNet, inPop, outAct) # model 1

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") #model 2

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, sameX, interaction1 = "departm_RS") #model 3

@@

#fsave(Model4, file ="estimated_M4")
